Propositions as Sessions, Mechanically

### 2015-04-27 paper/presentation.tex
Added a CP Example (not a translation from GV however).

\usepackage[latin1]{inputenc}
\usepackage{cmll}
\usepackage{lstmacros}
\usepackage{graphicx}
\usepackage{array}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{etoolbox}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usepackage[style=alphabetic,natbib=true]{biblatex}

@@

% GV/CP example listing details
\lstdefinelanguage{GV}{
    mathescape=true,
    morekeywords=[1]{case, connect, in, let, of, receive, select, send,
      terminate, to, with},
    sensitive=false,
    morecomment=[l]{//},
    morecomment=[s]{/*}{*/},
    morestring=[b]",
    basicstyle=\HappyMed,
    keywordstyle=[1]{\color{magenta}},
}

\lstnewenvironment{gv}{\lstset{language=GV}}{}

@@

\begin{frame}
\ftitle{Session Types}{1.5}
\begin{math}
  \begin{array}{cccc}
    S & \bnf  & \inpt{T}{S} & \mbox{receive value of type T, continue as S}\\
      & \altt & \outpt{T}{S} & \mbox{send value of type T, continue as S}\\
      & \altt & \Branch{l_i}{S_i}{i\in I} &
                           \mbox{offer between $|I|$ alternatives}\\
      & \altt & \Choice{l_i}{S_i}{i\in I} &
                           \mbox{choose between $|I|$ alternatives}\\
      & \altt & \mbox{\lstinline{end}} & \mbox{terminate session}\\
 \dual{S} & \bnf & \mbox{dual of S}\\
 \dual{(\inpt{T}{S})} & \bnf & \outpt{T}{\dual{S}}\\
 \dual{\Branch{l_i}{S_i}{i\in I}} & \bnf & \Choice{l_i}{\dual{S_i}}{i\in I}\\
            \vdots                & \cdots & \vdots
 \end{array}
\end{math}
\note{
one approach to achieve such correctness guarantees is by using binary session
types. [INTRODUCE BINARY SESSION TYPES HERE]. specifying a communication
protocol then involves providing mutual dual types to the communicating
parties.
}
\end{frame}

@@

@@

\begin{frame}
\frametitle{Motivation}
Informal proofs $=$ Error-prone $+$ Difficult to update
\note{
there has been a lot of activity in this area producing a variety of different
session-based type systems. many systems build on top of previous work adding
more complex features such as asynchronous, polymorphic and recursive session
types. extending informal proofs is error-prone and difficult because it is
not always clear which parts of a system's metatheory is affected by a
change. a formalised system would reduce likelihood of errors in the proofs
and aid extension since the theorem prover would highlight changes to the
metatheory
}
\end{frame}

\begin{frame}
\frametitle{CP Propositions}
\begin{math}
\begin{array}{rclc}
    \tp{A,B,C} & \bnf
          & \tp{A \otimes B} & \key{times: output A then behave as B}\\
          & \altt & \tp{A \parr B} & \key{par: input A then behave as B}\\
          & \altt & \cdots
\end{array}
\end{math}

For all propositions $A$, $\dualbt{(\dualbt{A})} = A$ (negation is an
involution)

\note{
Here are some exmaples of linear logic connectives with session type
interpretations. The operator times corresponds to output and the operator par
corresponds to input. Since it is based on the classical variant of linear
logic we have that negation is an involution.
}
\end{frame}

>>>>

\usepackage[latin1]{inputenc}
\usepackage{cmll}
\usepackage{lstmacros}
\usepackage{graphicx}
\usepackage{array}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{etoolbox}
\usepackage[cal=cm,calscaled=.80]{mathalfa}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usepackage[style=alphabetic,natbib=true]{biblatex}

@@

\newcommand\CPFSize{\fontsize{7}{8.4}\selectfont}
% CP/GV example listing details
\lstdefinelanguage{CP}{
    mathescape=true,
    morekeywords=[1]{case, inl, inr},
    sensitive=false,
    morecomment=[l]{//},
    morecomment=[s]{/*}{*/},
    morestring=[b]",
    basicstyle=\CPFSize,
    keywordstyle=[1]{\color{red}},
}
\lstdefinelanguage{GV}{
    mathescape=true,
    morekeywords=[1]{case, connect, in, let, of, receive, select, send,
      terminate, to, with},
    sensitive=false,
    morecomment=[l]{//},
    morecomment=[s]{/*}{*/},
    morestring=[b]",
    basicstyle=\HappyMed,
    keywordstyle=[1]{\color{magenta}},
}

\lstnewenvironment{cp}{\lstset{language=CP}}{}
\lstnewenvironment{gv}{\lstset{language=GV}}{}

@@

\begin{frame}
\ftitle{Session Types}{1.5}
\begin{math}
  \begin{array}{cccc}
    S & \bnf  & \inpt{T}{S} & \mbox{receive value of type T, continue as S}\\
      & \altt & \outpt{T}{S} & \mbox{send value of type T, continue as S}\\
      & \altt & \Branch{\mathcal{L}_i}{S_i}{i\in I} &
                           \mbox{offer between $|I|$ alternatives}\\
      & \altt & \Choice{\mathcal{L}_i}{S_i}{i\in I} &
                           \mbox{choose between $|I|$ alternatives}\\
      & \altt & \mbox{\lstinline{end}} & \mbox{terminate session}\\
 \dual{S} & \bnf & \mbox{dual of S}\\
 \dual{(\inpt{T}{S})} & \bnf & \outpt{T}{\dual{S}}\\
 \dual{\Branch{\mathcal{L}_i}{S_i}{i\in I}} & \bnf &
       \Choice{\mathcal{L}_i}{\dual{S_i}}{i\in I}\\
            \vdots                & \cdots & \vdots
 \end{array}
\end{math}
\note{
one approach to achieve such correctness guarantees is by using binary session
types. [INTRODUCE BINARY SESSION TYPES HERE]. specifying a communication
protocol then involves providing mutual dual types to the communicating
parties.
}
\end{frame}

@@

\begin{frame}[fragile]
\ftitle{CP Example}{0.5}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} cc}
Process A ($A_x$) & Process B ($B_x$)\\
$\colored\tp{Int \parr (String \otimes \1)}$
  ($\colored\tp{\dualbt{S}}$) &
$\colored\tp{Int \otimes (String \parr \bot)}$ ($\colored\tp{S}$)\\
\begin{cp}
$\colored\tm{x(y).x[w].(\mbox{put-thanks}_w \mid x[\,].0)}$
\end{cp} &
\begin{cp}
$\colored\tm{x[y].(\mbox{put-gift}_y \mid x(w).x().\mbox{get-response}_w)}$
\end{cp}
\end{tabular*}
\[\colored
\inference{
  \tm{B_x} \tpvdash \tp{\Gamma \comma \tmof{x}S}
  &
  \tm{A_x} \tpvdash \tp{\Delta \comma \tmof{x}\dualbt{S}}
}{Cut}{
  \tm{\nu x \of{S}.(B_x \mid A_x)} \tpvdash \tp{\Gamma \comma \Delta}
}
\]
\CPFSize
\colored
\begin{align*}
  \tm{\nu x.(B_x \mid A_x)} & =
    \tm{\nu x.(x[y].(\mbox{put-gift}_y \mid x(w).x().\mbox{get-response}_w)
          \mid x(y).x[w].(\mbox{put-thanks}_w \mid x[\,].0))}
    \\
                            & \becomes
    \tm{\nu y.(\mbox{put-gift}_y \mid
               \nu x.(x(w).x().\mbox{get-response}_w) \mid
                      x[w].(\mbox{put-thanks}_w \mid x[\,].0))}
    \\
                            & \becomes
    \tm{\nu y.(\mbox{put-gift}_y \mid
               \nu w.(\mbox{put-thanks}_w \mid
                      \nu x.(x().\mbox{get-response}_w \mid x[\,].0)))}
    \\
                            & \becomes
    \tm{\nu y.(\mbox{put-gift}_y \mid
               \nu w.(\mbox{put-thanks}_w \mid \mbox{get-response}_w))}
\end{align*}
\note{
Here are some exmaples of linear logic connectives with session type
interpretations. The operator times corresponds to output and the operator par
corresponds to input. Since it is based on the classical variant of linear
logic we have that negation is an involution.
}
\end{frame}

@@

\begin{frame}
\frametitle{Motivation}
Informal proofs $=$ Error-prone $+$ Difficult to update
\begin{itemize}
\item Subject Reduction: If $m : T$ and $m \rightarrow m'$ then $m' : T$
\end{itemize}
\note{
there has been a lot of activity in this area producing a variety of different
session-based type systems. many systems build on top of previous work adding
more complex features such as asynchronous, polymorphic and recursive session
types. extending informal proofs is error-prone and difficult because it is
not always clear which parts of a system's metatheory is affected by a
change. a formalised system would reduce likelihood of errors in the proofs
and aid extension since the theorem prover would highlight changes to the
metatheory
}
\end{frame}
